Analysis: Machine Learning - ReqAmt

### 2021-12-12 archive/analysis-ML-RA-old.Rmd
ML update - new models, started outliers

---
title: "Analysis: Machine Learning - ReqAmt - Outliers"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---

@@

## Load Data
Load the analysis data from the `processed_data` folder in the project file.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","analysisdata.rds")

@@

* Drop the unused variables
* Perform feature engineering
* Transform Incident Month into ordinal variable
* Remove outlier
* Set the random seed to 123
* Split the dataset into 70% training, 30% testing, stratified on incidentType
* 5-fold cross validation, 5 times repeated, stratified on incidentType for the CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#requested amount filter
req_data <- analysisdata %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, ResponseDays, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, ReqAmt)

#transform incident month into ordinal variable
req_data$IncidentMonth <- ordered(req_data$IncidentMonth, levels = 1:12,
                                  labels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))

#not enough values per state, so create a variable that is FEMA Regions
req_data$FEMARegion <- ifelse((req_data$state == "CT" |
                                 req_data$state == "ME" |
                                 req_data$state == "MA" |
                                 req_data$state == "NH" |
                                 req_data$state == "RI" |
                                 req_data$state == "VT"),
                              "Region I",
                              ifelse((req_data$state == "NJ" |
                                         req_data$state == "NY" |
                                         req_data$state == "PR" |
                                         req_data$state == "VI"),
                                      "Region II",
                                       ifelse((req_data$state == "DE" |
                                                 req_data$state == "MD" |
                                                 req_data$state == "PA" |
                                                 req_data$state == "VA" |
                                                 req_data$state == "DC" |
                                                 req_data$state == "WV"),
                                              "Region III",
                                               ifelse((req_data$state == "AL" |
                                                         req_data$state == "FL" |
                                                         req_data$state == "GA" |
                                                         req_data$state == "KY" |
                                                         req_data$state == "MS" |
                                                         req_data$state == "NC" |
                                                         req_data$state == "SC" |
                                                         req_data$state == "TN"),
                                                      "Region IV",
                                                      ifelse((req_data$state == "IL" |
                                                               req_data$state == "IN" |
                                                               req_data$state == "MI" |
                                                               req_data$state == "MN" |
                                                               req_data$state == "OH" |
                                                               req_data$state == "WI"),
                                                            "Region V",
                                                             ifelse((req_data$state == "AR" |
                                                                         req_data$state == "LA" |
                                                                         req_data$state == "NM" |
                                                                         req_data$state == "OK" |
                                                                         req_data$state == "TX"),
                                                                      "Region VI",
                                                                       ifelse((req_data$state == "IA" |
                                                                                 req_data$state == "KS" |
                                                                                 req_data$state == "MO" |
                                                                                 req_data$state == "NE"),
                                                                              "Region VII",
                                                                               ifelse((req_data$state == "CO" |
                                                                                         req_data$state == "MT" |
                                                                                         req_data$state == "ND" |
                                                                                         req_data$state == "SD" |
                                                                                         req_data$state == "UT" |
                                                                                         req_data$state == "WY"),
                                                                                      "Region VIII",
                                                                                       ifelse((req_data$state == "AZ" |
                                                                                                 req_data$state == "CA" |
                                                                                                 req_data$state == "HI" |
                                                                                                 req_data$state == "NV"),
                                                                                              "Region IX",
                                                                                               ifelse((req_data$state == "AK" |
                                                                                                         req_data$state == "ID" |
                                                                                                         req_data$state == "OR" |
                                                                                                         req_data$state == "WA"),
                                                                                                      "Region X",
                                                                                                      NA
                                                                                                      )
                                                                                              )
                                                                                      )
                                                                              )
                                                                      )
                                                            )
                                                      )
                                              )
                                     )
                              )

#drop state variable
req_data <- req_data %>%
              dplyr::select(-c(state))

#add to "other" category for incident Types that are infrequent
req_data$incidentType[(req_data$incidentType == "Chemical" |
                        req_data$incidentType == "Coastal Storm" |
                        req_data$incidentType == "Dam/Levee Break" |
                        req_data$incidentType == "Mud/Landslide" |
                        req_data$incidentType == "Volcano" |
                        req_data$incidentType == "Terrorist" |
                        req_data$incidentType == "Snow")] <- "Other"

#remove Hurricane Maria outlier
req_data <-subset(req_data, ReqAmt< 4000000000)

#set random seed
set.seed(123)

@@

#use the recipe specified earlier
```
---

@@

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
BT_res <- BT_wflow %>%
          tune::tune_grid(resamples = req_data_bootsplit,
                          grid = tree_grid,
                          metrics = metric_set(rmse),
                          control = control_resamples(save_pred = TRUE,
                                                      extract = function(x) extract_model(x)))

@@

### 6. Model evaluation
```{r}
#bag roots function to extract most important variables
bag_roots <-  function(x){
  x %>% 
    dplyr::select(.extracts) %>% 
    unnest(cols = c(.extracts)) %>% 
    dplyr::mutate(models = map(.extracts,
                               ~.x$model_df)) %>% 
    dplyr::select(-.extracts) %>% 
    unnest(cols = c(models)) %>% 
    mutate(root = map_chr(model,
                          ~as.character(.x$fit$frame[1, 1]))) %>%
    dplyr::select(root)  
}

BT_IV <- bag_roots(BT_res) %>% 
            ggplot(mapping = aes(x = fct_rev(fct_infreq(root)))) + 
            geom_bar() + 
            coord_flip() + 
            labs(x = "root", y = "count")
BT_IV

#save file
BT_IV_file = here("results","BT-IV-outlier.png")
ggsave(filename = BT_IV_file, plot = BT_IV)

>>>>

---
title: "Analysis: Machine Learning - ReqAmt"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---

@@

## Load Data
Load the ReqAmt analysis data from the `processed_data` folder in the project file.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","analysisdata-RA.rds")

@@

* Set the random seed to 123
* Split the dataset into 70% training, 30% testing, stratified on incidentType
* 5-fold cross validation, 5 times repeated, stratified on incidentType for the CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#set random seed
set.seed(123)

@@

#use the recipe specified earlier (line 133)
```
---

@@

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
BT_res <- baguette::bag_tree(cost_complexity = tune(),
                                   tree_depth = tune(),
                                   min_n = tune()) %>%
          parsnip::set_engine("rpart",
                              times = 10) %>%
          parsnip::set_mode("regression") %>%
          tune::tune_grid(preprocessor = BT_rec,
                          resamples = req_data_bootsplit,
                          grid = tree_grid,
                          metrics = metric_set(rmse))

@@

### 6. Model evaluation
```{r}
#extract model from final fit
x_BT <- best_BT_fit$fit$fit$fit

#vip package doesn't support BT
